Development of a 3D display

### 2016-05-14 paper.tex
Added & finished chapter 3

\begin{figure}[h]
\vspace{4mm}
\includegraphics[width=\textwidth]{./images/board.png}
\vspace{-4mm}
\caption{The PCB layout in Eagle. Red = upper copper layer, blue = lower copper layer,
green = vias, grey = silkscreen.} 
\vspace{6mm}
\end{figure}

@@

>>>>

\begin{figure}[h]
\vspace{4mm}
\includegraphics[width=\textwidth]{./images/board-layout.png}
\vspace{-4mm}
\caption{The PCB layout in Eagle. Red = upper copper layer, blue = lower copper layer,
green = vias, grey = silkscreen.} 
\vspace{6mm}
\end{figure}

@@

\chapter{Fabrication}

\section{Circuit Boards}

\subsection{Fabrication}

There's no shortage of firms who produce PCBs according to a custom layout. One of the cheapest 
providers to ship small production runs within a reasonable timeframe and with good production
quality turned out to be OSHPark. They only produce in multiples of 3, so I ordered 12 instead of
10 (which I'd have done anyway to have some spare boards). I submitted my order on the 5th of 
October 2014 and paid US\$ 47. Ten days later I received a message that the boards were finished,
and after another ten days of waiting they arrived at my doorstep.

\begin{figure}[h]
\vspace{3mm}
\includegraphics[width=\textwidth]{./images/board-finished.jpeg}
\vspace{-8mm}
\caption{A finished circuit board} 
\end{figure}

\subsection{Assembly}

As soon as I got the boards in the mail, I started soldering the parts to them. I started with the
LEDs to avoid having to solder in the tight gap that would later be formed between the LEDs and
their driver chips. It took me two hours to solder the first board, but with increasing practice
the time for one row of LEDs got down to 25 or 30 minutes.

The chips were a lot easier to solder than the LEDs, despite the distance between the individual
pins only being \SI{0.635}{\milli\meter}. I could simply drown all the pins in solder, and then 
wick away all the solder bridges. That way I could solder twelve pins within the blink of an eye.

\subsection {Testing and Debugging}

With the boards finished, I wanted to know whether or not they work. In order to find out, I wrote
a little program on the microcontroller to light the LEDs up in all possible colours consecutively.
It did not work at the first try, so I used an oscilloscope to see if data was being sent at all. 
This was the case, the SPI library on the Teensy performed its task flawlessly.

It turned out that the culprit were the bad contacts. Instead of soldering the wires onto the board
and the microcontroller, I simply inserted them into the respective holes. This caused the 
connection to have too much parasitic capacitance, and the signals (which were being sent at 4 MHz)
did not have time to propagate up to their destination.

I soldered the cables on properly and sure enough, the board worked as intended. I repeated this
procedure for all ten boards to see if they all worked as well as the first one, which wasn't the 
case.

Because of reasons unbeknownst to me, one particular LED driver started turning on all of its LEDs
at a moderate brightness as soon as I powered it up --- even though no data was being sent at all.
At first I thought that the serial data input had a solder bridge pulling it up to \vdd, and
immediately sending a whole lot of 1's upon powering up the board. However, the data input pin was
not close enough to a positive supply line for this to be an issue. Additionally, there would have 
to be a clock signal as well as a pulse on the latch pin for the LEDs to light up. I gave up and 
assumed that the chip was faulty, and when I replaced it with a spare part everything worked 
perfectly.

Apart from that, there were no major issues. I had a few LEDs which didn't light up when they 
should have. One of them had a solder bridge between the common anode and the cathode of the red
LED, which made a voltage drop across the red LED impossible. Two were simply soldered on poorly
and weren't connected to the drivers at all. A few other LEDs didn't work and I found no obvious
issues. I replaced them with spares, which did the trick.

\section{Physical Assembly}

\subsection{Frame}

The frame was built out of beech wood strips with a square cross section and a side length of 
\SI{1.5}{\centi\meter}. The baseplate is made out of plywood and has the dimensions 
\SI{15}{\centi\meter} by \SI{15}{\centi\meter} by \SI{1}{\centi\meter}. On it, I mounted two 
vertical wood strips with a length of \SI{20}{\centi\meter}, which were both stabilise by three
struts. A horizontal board, mounted to the vertical strips at a height of \SI{20}{\centi\meter}, 
holds the motor.

On the top of the two vertical strips I placed another strip to hold the shaft in its position. 
A simple hole served as a bearing, which I later replaced with a plastic insert (kindly sponsored
by IGUS).

\subsection{Slip Rings}

In order to supply the rotating parts with electricity, I mounted a wooden disc right at the bottom
of the shaft. On both sides of that disc I glued on a round piece of copper sheet, which I then
connected to cables going up to where the electronics would later be.

On the stationary end, two long pieces of copper sheet serve as contacts. Because of their 
spring-like properties, they exert a certain amount of pressure on the rotating disc.

Despite this, a considerable amount of pressure was necessary to form a good contact. The contact
pressure proved to be a primitive, but usable way of roughly adjusting the rotational speed of
the display. Sadly it's quite an inefficient approach: The motor uses a lot of current when 
friction is applied to its load. It becomes noticeably warm after a few minutes of use, and 
measurements on the current limiting resistors have shown that it draws about \SI{6.5}{\ampere} at
\SI{5}{\volt}.

This constant pressure, in addition to the fact that copper is quite a soft metal, results in 
considerable wear on the slip rings. Luckily, the stationary ends can be replaced without much 
hassle. The rotating part would be a pain to replace, but since wear is distributed across the 
whole copper disc, it will be a long time before this needs to be done.

\subsection{Controller}

The microcontroller is mounted on another wooden disc, located a few centimetres above the slip
rings. It also hosts the hall sensor, which is used to measure the rotational speed and calculate
the position of the rotating part at any given moment. The corresponding magnet is mounted on a 
piece of threaded rod, which itself is mounted on the same board as the motor.

Additionally, the \SI{2200}{\micro\farad} capacitor is located on the upper wooden disc. In the 
end, this place has become a bit of a mess, especially with all the cables going up to the circuit 
boards. Luckily this isn't visible while the apparatus is running.

\subsection{Circuit Boards}

The boards were already equipped with components, but what was missing was the hold to mount them
on the shaft. The diameter of this shaft is \SI{4}{\milli\meter}. In order to find a good hole 
size, I drilled a few test holes on a spare board. A hole of \SI{4}{\milli\meter} was so small for
the boards to fit on the shaft, while \SI{4.5}{\milli\meter} was obviously too big. After 
chamfering the top end of the shaft, the board with a \SI{4}{\milli\meter} hole just barely could
slide on the shaft. This tight fit proved to be advantageous, because torque can now be transmitted
reliably from the shaft to the boards, and they don't slide around.

\chapter{Programming}
